\begin{document}

\frontmatter

\title{Managing Complex Water Resource Systems for Ecological Integrity: Evaluating Tradeoffs and Uncertainty}
\author{Ryan Richard Morrison}
\degreesubject{Ph.D., Civil Engineering}
\degree{Philosophy of Science \\ Engineering}
\documenttype{Dissertation}
\previousdegrees{B.S., Civil Engineering, Washington State University, 2003
\\M.S., Civil Engineering, Washington State University, 2006}
\date{May, \thisyear} % leave blank for today's date
\maketitle
\makecopyright

\begin{dedication}
   To my parents for their support and encouragement. Aren't you glad I'm not an ornithologist? \\[3ex]
\end{dedication}

\begin{acknowledgments}
   \vspace{1.1in}
   I would like to thank my friend and advisor, Mark Stone, for his trust and encouragement during my PhD program. I would not have moved to New Mexico without knowing you would be an excellent mentor and beer-drinking companion.

   \noindent{I  am also gracious for the support and free whitewater boating trips provided by the great New Mexico water resources community, including Mindy Harm Benson, Mike Harvey, Reed Benson, Dagmar Llewellyn, Steve Harris, and others.}

    \noindent{In addition, I want to thank my committee members, Andy Schuler, Mindy Harm Benson, and Vanessa Valentin, for their advice and guidance during my research.}

   \noindent{Finally, I still can't believe the encouragement given by my wife, Leslie, even while we lived 1,000 miles apart for two years. Thank you, Sweetie.\\ \emph{``Home is wherever I'm with you.''}}
\end{acknowledgments}

\maketitleabstract %(required even though there's no abstract title anymore)

\begin{abstract}
   Water resource systems often contain numerous components that are intertwined or contradictory, such as power production, water delivery, recreation, and environmental needs. This complexity makes it difficult to holistically assess management alternatives. In addition, hydroclimatic and ecological uncertainties complicate efforts to evaluate the impacts of management scenarios. We need new tools that are able to inform managers and researchers of the tradeoffs or consequences associated with flow alternatives, while also explicitly incorporating sources of uncertainty. My research addresses this limitation using two modeling approaches: stochastic system dynamics modeling and Bayesian network modeling. I developed a stochastic system dynamics model to evaluate the impacts of environmental flow alternatives on multiple water users in the Rio Chama basin, New Mexico. Specifically, my work examined the influence of flow alternatives on cottonwood recruitment, reservoir storage, hydropower production, and whitewater boating. In addition, I coupled two-dimensional hydrodynamic and Bayesian network models to assess the impacts of management scenarios on cottonwood recruitment on the Gila River, New Mexico. The Bayesian network approach explicitly incorporated spatial variability, as well as hydrologic and ecological uncertainties. These methods are useful for more thoroughly assessing the tradeoffs of management decisions, integrating system components within a holistic framework, and evaluating ecological consequences of management scenarios at fine spatial scales.
\clearpage %(required for 1-page abstract)
\end{abstract}

\tableofcontents
\listoffigures
\listoftables

%\begin{glossary}{Longest  string}
%   \item[$a_{lm}$]
%      Taylor series coefficients, where $l,m = \{0..2\}$
%   \item[$A_{\bf{p}}$]
%      Complex-valued scalar denoting the amplitude and phase.
%   \item[$A^T$]
%      Transpose of some relativity matrix.
%\end{glossary}

\mainmatter


\setlength{\parskip}{0 cm} % Set the spacing between paragraphs
\setlength{\parindent}{1cm} % Remove all paragraph indentation
\setlength{\overfullrule}{0pt} % Prevent black boxes from appearing at end of lines

\lefthyphenmin4
\righthyphenmin4

%\pretolerance=10000
%\tolerance=2000
%\emergencystretch=10pt

\linespread{1.3} % A line spread factor of 1.3 produces 1.5 lines spacing (not very intuitive, I know)

\raggedbottom % Allow page bottoms to be empty so the paragraphs aren't stretched across the page at the end of sections